---
title: "RepData_PeerAssessment1"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## GitHub Documents

This is an R Markdown format used for publishing markdown documents to GitHub. When you click the **Knit** button all R code chunks are run and a markdown file (.md) suitable for publishing to GitHub is generated.

## Including Code

You can include R code in the document as follows:

```{r activity, echo=TRUE }
library(readr)
activity <- read_csv("Reproducible Research/week2/activity.csv")
#View(activity)
```
## preprocessing the data
```{r pactivity, echo=TRUE}
# Check for missing values
sum(is.na(activity$steps))

# Replace missing values if necessary (using mean for the interval)
activity_clean <- activity
activity_clean$steps[is.na(activity_clean$steps)] <- 
  ave(activity_clean$steps, activity_clean$interval, FUN = function(x) mean(x, na.rm = TRUE))


```
## calculating total steps per day

```{r tactivity_clean, echo=TRUE }
total_steps_per_day <- aggregate(steps ~ date, data = activity_clean, sum)

```
## Including Plots

You can also embed plots, for example:

```{r Hactivity_clean, echo=TRUE}
hist(total_steps_per_day$steps, 
     main = "Total Steps per Day", 
     xlab = "Steps", 
     col = "blue")

```
## calculate mean and median steps per day

```{r mactivity_clean, echo=TRUE }
mean_steps <- mean(total_steps_per_day$steps)
median_steps <- median(total_steps_per_day$steps)
```
## average daily activity pattern
```{r activity_clean, echo=TRUE}
avg_steps_interval <- aggregate(steps ~ interval, data = activity_clean, mean)

# Plot the time series
plot(avg_steps_interval$interval, avg_steps_interval$steps, type = "l", 
     main = "Average Daily Activity Pattern", xlab = "Interval", ylab = "Average Steps")

```
## Differences Between Weekdays and Weekends

```{r Dactivity_clean, echo=TRUE }
activity_clean$day_type <- ifelse(weekdays(as.Date(activity_clean$date)) %in% c("Saturday", "Sunday"),   "Weekend", "Weekday")

```

## Analyze activity patterns for weekdays vs. weekends
```{r Wactivity_clean, echo=TRUE}
library(ggplot2)
avg_steps_day_type <- aggregate(steps ~ interval + day_type, data = activity_clean, mean)

# Plot using ggplot2
ggplot(avg_steps_day_type, aes(x = interval, y = steps, color = day_type)) +
  geom_line() +
  labs(title = "Activity Patterns: Weekdays vs Weekends", x = "Interval", y = "Average Steps") +
  theme_minimal()

```
Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
